---
title: "JSM Panel Count"
author: "Steve Simon"
date: "7/29/2019"
output: html_document
---

The text file, panels.txt, was obtained from the online program of JSM2019. I went to the advanced search tab, clicked on the panels option to get only panel discussions, and cut and pasted the resulting output into a text file.

```{r read}
suppressWarnings(suppressMessages(library(gender)))
suppressWarnings(suppressMessages(library(magrittr)))
suppressWarnings(suppressMessages(library(tidyverse)))
raw_text <- readLines("panels.txt")
```

The text file has a consistent format that makes it easy to identify the first panelist is on a line starting with the text "Panelists:\t" and the last panelist appears just before the line with the text "Floor Discussion".

```{r extract}
i <- grep("^Panelists:\t", raw_text)
j <- grep("Floor Discussion", raw_text)
for (k in 1:length(i)) {
  cat(paste(raw_text[i[k]:j[k]], collapse="\n"))
  cat("\n\n")
}
full_name <- NULL
panel <- NULL
for (k in 1:length(i)) {
  full_name <- c(full_name, raw_text[i[k]:(j[k]-1)])
  panel <- c(panel, rep(k, j[k]-i[k]))
}
full_name %>%
  sub("^Panelists:\t", "", .) %>%
  sub(" .*", "", .) -> first_name
```

The gender package takes a first name and compares it to a database of names from Social Security that calculates the proportion of males and of females in that database who have a specific first name. Names like "Mary" and "John" will belong almost exclusively to one gender, but names like "Pat" will be more evenly split.

```{r gender}
identify_gender <- gender(first_name)
identify_gender %>%
  distinct(name, .keep_all=TRUE) %>%
  full_join(data.frame(panel=panel, name=first_name, stringsAsFactors=FALSE)) -> panel_demographics
```

The gender package will only report results on names that are found in their database, and this would exclude some (but not all) "foreign" names. Any of these cases, plus any cases where the gender package rates the probability of male/female somewhere in the ambiguous middle (between 20% and 80%) will be coded as unknown/ambiguous.

Here are the names with unknown gender. You could probably resolve some of these easily.

```{r count-unknown}
unknown_gender <- is.na(panel_demographics$gender)
panel_demographics %>%
  filter(unknown_gender) %>%
  pull(name) %>%
  unique %>%
  sort %>%
  paste(collapse=", ") %>%
  print
```

Here are the ambiguous names.

```{r count-ambiguous}
ambiguous_gender <- panel_demographics$proportion_male > 0.20 & panel_demographics$proportion_male < 0.80
panel_demographics %>%
  filter(ambiguous_gender) %>%
  pull(name) %>%
  unique %>%
  sort %>%
  paste(collapse=", ") %>%
  print
```

Code both the unknown and ambiguous names into the same category.

```{r code-unknown-and-ambiguous}
panel_demographics$gender[unknown_gender] <- "unknown/ambiguous"
panel_demographics$gender[ambiguous_gender] <- "unknown/ambiguous"
table(panel_demographics$gender)
```

It is very easy to get a count of female and male names by panel. The proportion of females excludes any names that were not cdoed as unknown/ambiguous.

```{r summarize}
panel_demographics %<>% arrange(panel, gender)
n_panels <- max(panel_demographics$panel)
name_list <- rep("", n_panels)
for (i in 1:n_panels) {
  name_list[i] <- paste(panel_demographics$name[panel_demographics$panel==i], collapse=", ")
}
panel_demographics %>%
  mutate(f=as.numeric(gender=="female")) %>%
  mutate(m=as.numeric(gender=="male")) %>%
  mutate(u=as.numeric(gender=="unknown/ambiguous")) %>%
  group_by(panel) %>%
  summarize(
    f=sum(f),
    m=sum(m),
    u=sum(u)) %>%
  mutate(p_f=round(f/(f+m), 2)) %>%
  bind_cols(tibble(name_list)) %>%
  arrange(p_f) %>%
  data.frame -> panel_summary
panel_summary %>%
  summarize_at(c("f", "m", "u"), sum) %>%
  mutate(p_f=round(f/(f+m), 2)) -> panel_summary_total_row
panel_summary         
panel_summary_total_row
```

Save everything for possible re-use.

```{r save}
save.image(file="panel-count.RData")
```